Mutation testing: Clustering the mutants

### 2021-11-29 chapters/2_introduction.tex
Added introduction + selecting mutation

\documentclass[../main]{subfiles}
\begin{document}

\chapter{Introduction}
\label{ch:introduction}
Software testing is a critical part in the software engineering process. 
It is a standardized in the industry and is used for quality assurance\cite{ISOtesting}. 
Software testing can detect software bugs during the development process and can also serve for regression purposes\cite{Bertolino2007SoftwareDreams}. 
During the software testing process developers write automated tests called unit tests. 
These tests can be run by testing frameworks such as JUnit and TestNG\cite{JUnit, TestNG}.
Monitoring the quality of test code has shown that tests with lower quality leads to more defect-prone production code\cite{Spadini2018OnQuality}. 
Research has also found that if the production code grows, the test code grows along with it\cite{Zaidman2008MiningCode}. 
Quality control on test suites is therefore important as the maintenance on tests can be difficult and generates risks if done incorrectly\cite{Horvath2015TestMetrics}.
\newline
The growing complexity and size of code bases consequently lead to incomprehensible tests.
As a result test bugs can occur. 
A test bug is a test that fail even tough the program is correct(\textit{false positive}). 
Another bug, which is even worse, are tests that do not fail when the program is not working as desired(\textit{false negative}).
The \textit{false negative} is a problem when breaking changes are not detected by the test suite.
\newline
To solve this issue we can measure the fault detecting capability of a test suite, this is called the test suite effectiveness.
Test suite effectiveness is measured by the number of faulty versions of a \acrfull{sut} that are detected by a test suite.
Real faults are unknown in advance, we can use mutation testing as a proxy measurement. 
Research has found statistically significant evidence which correlates mutant detection with real fault detection\cite{Just2014AreTesting}.
\newline
\todo[inline]{is this your approach? the current general approach?}
Mutation testing tools create small faulty versions of the program.
The test suite is then run to verify if these faults are detected.
These faults are called mutants and are created by mutators.
Mutators mutate specific expressions or statements in the source code. 
To prevent changing the overall functionality of the program, mutants are represented as very small changes\cite{Lipton1978HintsProgrammer}.
\newline
After all the mutants are executed the mutation score can be calculated. 
The mutation score is the percentage of mutants killed divided by the total amount of mutants executed.
If a large amount of mutants survives, it might be an indication that the quality of the test suite is poor as programming errors remain undetected.
Mutation testing has considerable drawbacks\cite{Pizzoleto2019}, such as the equivalent mutant problem and being resource expensive.

\section{Problem statement}
Mutation testing requires a significant amount of computational resources\cite{Kintis2016EffectiveMutation}. 
The high cost requirement is often a barrier for adopting mutation testing \cite{Pizzoleto2019}.
A lot of techniques and methods have been developed to improve the performance, however, most of these approaches are not as effective as mutation testing a full set of mutants\cite{Pizzoleto2019,Yao2014}. 
\newline
A technique to reduce the number of mutants or the number of test case executions is mutation clustering.
Mutation clustering aims to reduce the amount of mutants to be executed by clustering mutants\cite{Ma2016,Yu2019PossibilityScope}.
Clustering is a technique used for the grouping of data, for pattern recognition and classification, based on some similar characteristics\cite{Alsabti1997}.
The clustering of mutants has been researched with promising results\cite{Ji2009,Wilinski2015,Ma2016}. For example Ma et al., \cite{Ma2016} clustered the mutants for expressions. 
However in their paper they indicate that the choosing expressions as the domain of their cluster is a limitation.
A recurring problem in the research of mutation clustering is to determine the means of domains as the centre of a cluster\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
\newline
To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
Our goal is to cluster any mutant that is generated within the Java programming language.
This research aims to remove the scoping limitations present in existing research.
There should be no requirements for which mutants can be clustered.
We will do this with two different approaches: qualitative and quantitative. 
We structure our research on the following research questions:
\newline
\textbf{Research Question 1}: What set of characteristics can we identify for clustering generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
\textbf{Research Question 1.1}: How do the existing mutation testing tools for Java compare to each other?
\newline
\textbf{Research Question 2}: How can we train a machine learning model to recognize and cluster generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
We start with comparing existing mutation testing tools. 
We will use the existing literature as a starting point and research what would be the best tool to use for this thesis.
\newline
After we have selected a tool we will research existing clustering techniques.
Based on the results of this research we select and implement a clustering algorithm that will cluster one ore more characteristics of the mutants. 
We will randomly select on mutant per of each cluster to be executed by the mutation tooling.
We will then compare the mutation score of the clustered mutants with the mutation score of a full set of mutants.
\newline
To answer our second research question we will train a machine learning model to cluster mutants.
We can then repeat the comparison we did for research question one. 
\newline
We want to see how the algorithm/model performs on big and small projects.
These project must also require a test suite.
\newline
We chose to focus on projects that use Java as main programming language because it is one of the most popular languages\cite{Tiobe}.

>>>>

wel e\documentclass[../main]{subfiles}
\begin{document}

\chapter{Introduction}
\label{ch:introduction}
Software testing is a critical part in the software engineering process. 
It is a standardized in the industry and is used for quality assurance\cite{ISOtesting}. 
Software testing can detect software bugs during the development process and can also serve for regression purposes\cite{Bertolino2007SoftwareDreams}. 
During the software testing process developers write automated tests called unit tests. 
These tests can be run by testing frameworks such as JUnit and TestNG\cite{JUnit, TestNG}.
Monitoring the quality of test code has shown that tests with lower quality leads to more defect-prone production code\cite{Spadini2018OnQuality}. 
Research has also found that if the production code grows, the test code grows along with it\cite{Zaidman2008MiningCode}. 
Quality control on test suites is therefore important as the maintenance on tests can be difficult and generates risks if done incorrectly\cite{Horvath2015TestMetrics}.
\newline
The growing complexity and size of code bases consequently lead to incomprehensible tests.
As a result test bugs can occur. 
A test bug is a test that fail even tough the program is correct(\textit{false positive}). 
Another bug, which is even worse, are tests that do not fail when the program is not working as desired(\textit{false negative}).
The \textit{false negative} is a problem when breaking changes are not detected by the test suite.
\newline
To solve this issue we can measure the fault detecting capability of a test suite, this is called the test suite effectiveness.
Test suite effectiveness is measured by the number of faulty versions of a \acrfull{sut} that are detected by a test suite.
Real faults are unknown in advance, we can use mutation testing as a proxy measurement. 
Existing research has found statistically significant evidence which correlates mutant detection with real fault detection\cite{Just2014AreTesting}.
\newline
Mutation testing tools create small faulty versions of the program.
The test suite is then run to verify if these faults are detected.
These faults are called mutants and are created by mutators.
Mutators mutate specific expressions or statements in the source code. 
To prevent changing the overall functionality of the program, mutants are represented as very small changes\cite{Lipton1978HintsProgrammer}.
\newline
After all the mutants are executed the mutation score can be calculated. 
The mutation score is the percentage of mutants killed divided by the total amount of mutants executed.
If a large amount of mutants survives, it might be an indication that the quality of the test suite is poor as programming errors remain undetected.
Mutation testing has considerable drawbacks\cite{Pizzoleto2019}, such as the equivalent mutant problem and being resource expensive.

\section{Problem statement}
Mutation testing requires a significant amount of computational resources\cite{Kintis2016EffectiveMutation}. 
The high cost requirement is often a barrier for adopting mutation testing \cite{Pizzoleto2019}.
A lot of techniques and methods have been developed to improve the performance, however, most of these approaches are not as effective as mutation testing a full set of mutants\cite{Pizzoleto2019,Yao2014}. 
\newline
A technique to reduce the number of mutants or the number of test case executions is mutation clustering.
Mutation clustering aims to reduce the amount of mutants to be executed by clustering mutants\cite{Ma2016,Yu2019PossibilityScope}.
The clustering of mutants has been researched with promising results\cite{Ji2009,Wilinski2015,Ma2016}. For example Ma et al., \cite{Ma2016} clustered the mutants for expressions. 
However in their paper they indicate that the choosing expressions as the domain of their cluster is a limitation.
A recurring problem in the research of mutation clustering is to determine the means of domains as the centre of a cluster\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
\newline
To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
Our goal is to cluster any mutant that is generated within the Java programming language.
This research aims to remove the scoping limitations present in existing research.
There should be no requirements for which mutants can be clustered.
We do this with two different techniques: qualitative and quantitative.
We devise a white and black box approach for clustering mutants.
Our white box approach contains a qualitative analysis on mutants and a methodology to cluster them.
The black box approach makes use of a machine learning model to cluster mutants.
We structure our research on the following research questions:
\newline
\textbf{Research Question 1}: What set of characteristics can we identify for clustering generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
\textbf{Research Question 1.1}: How do the existing mutation testing tools for Java compare to each other?
\newline
\textbf{Research Question 2}: How can we train a machine learning model to recognize and cluster generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
We start with comparing existing mutation testing tools. 
We will use the existing literature as a starting point and research what would be the best tool to use for this thesis.
\newline
After we have selected a tool we will research existing clustering techniques.
Based on the results of this research we select and implement a clustering algorithm that will cluster one ore more characteristics of the mutants. 
We will randomly select on mutant per of each cluster to be executed by the mutation tooling.
We will then compare the mutation score of the clustered mutants with the mutation score of a full set of mutants.
\newline
To answer our second research question we will train a machine learning model to cluster mutants.
We can then repeat the comparison we did for research question one. 
\newline
We want to see how the algorithm/model performs on big and small projects.
These project must also require a test suite.
\newline
We chose to focus on projects that use Java as main programming language because it is one of the most popular languages\cite{Tiobe}.

### 2021-11-29 mutation-workshop/paper.tex
Added introduction + selecting mutation

\usepackage[backend=biber, style=ieee, citestyle=numeric-comp,uniquelist=false, maxcitenames=2, maxbibnames=9, mincitenames=1]{biblatex}
\bibliography{../refs/main}

@@

\and
\IEEEauthorblockN{Rasjaad Basarat}
\IEEEauthorblockA{\textit{Parallel Computing Systems} \\ %% TODO: CHANGE THIS
\textit{University of Amsterdam}\\
Amsterdam, the Netherlands \\
rasjaad@nivero.io}
\and
\IEEEauthorblockN{Thomas Biesaart}
\IEEEauthorblockA{\textit{Parallel Computing Systems} \\ %% TODO: CHANGE THIS
Amsterdam, the Netherlands \\
thomas@emailhere.com}
}

\maketitle
% \todo[inline]{FROM REVIEWER: Make sure to be consistent with singular/plural (e.g. mutation operators will result in mutants, not just a mutant)}
% \todo[inline]{FROM REVIEWER:  I don't quite follow the discussion of the execution time problem. After all, even in a decentralized approach I can run multiple approaches on the same hardware and thus make a fair comparison. I suppose the point you're trying to make is comparison *across multiple papers/studies"? Please clarify.}
\begin{abstract}
Mutation testing is a computationally expensive technique that requires the ability to run tests.
Different techniques have been proposed to reduce the cost of mutation testing.
One of these techniques is the clustering of mutants.
By clustering mutants we can execute less mutants to reduce the cost.
Different techniques have been proposed to reduce the cost of mutation testing.
One of these techniques is the clustering of mutants.
Our research consists of two parts in which we cluster mutants.
To cluster mutants we need a mutation testing tools.
Three mutation testing tools for Java are compared and one is selected to use for this research
The first part consists of a white box approach and the second part of a black box approach.
Our research uses the mutation testing tool PIT.
We identified characteristics to represent the mutants such that we can cluster them.
For our white box approach we used hierarchical clustering and for our black box approach we trained the fuzzy c-means  model.
We calculated a weighted mutation score and compared this with the mutation score of a full set of mutants executed.
Results show that with hierarchical clustering we can reduce the amount of mutants executed while maintaining the effectiveness.
The clusters generated by the machine learning model were less accurate and showed a significant decrease in effectiveness.
\end{abstract}

@@

\section{Introduction}
\subsection{Contributions}
\subsection{Outline}

@@

\section{Comparing mutation testing tooling}
\section{Qualitative clustering approach}
\section{Quantitative clustering approach}
\section{Experiment design}

\section{Discussion}

>>>>

\usepackage[backend=biber, style=ieee, citestyle=numeric-comp,uniquelist=false, maxcitenames=2, maxbibnames=9, mincitenames=1]{biblatex}
\bibliography{../main}
\bibliography{../custom}

%% uncomment this and change compiler for dyslexic version
%\usepackage{fontspec}
%\setmainfont{OpenDyslexic}

@@

\and
\IEEEauthorblockN{Rasjaad Basarat}
\IEEEauthorblockA{\textit{Parallel Computing Systems} \\ %% TODO: CHANGE THIS
\textit{University of Amsterdam}\\
Amsterdam, the Netherlands \\
rasjaad@nivero.io}
\and
\IEEEauthorblockN{Thomas Biesaart}
\IEEEauthorblockA{\textit{Parallel Computing Systems} \\ %% TODO: CHANGE THIS
Amsterdam, the Netherlands \\
thomas.biesaart@protonmail.com}
}

\maketitle
% \todo[inline]{FROM REVIEWER: Make sure to be consistent with singular/plural (e.g. mutation operators will result in mutants, not just a mutant)}
% \todo[inline]{FROM REVIEWER:  I don't quite follow the discussion of the execution time problem. After all, even in a decentralized approach I can run multiple approaches on the same hardware and thus make a fair comparison. I suppose the point you're trying to make is comparison *across multiple papers/studies"? Please clarify.}
\begin{abstract}
Mutation testing is a computationally expensive technique that requires the ability to run tests.
Different techniques have been proposed to reduce the cost of mutation testing.
One of these techniques is the clustering of mutants.
By clustering mutants we can execute less mutants to reduce the cost.
Different techniques have been proposed to reduce the cost of mutation testing.
One of these techniques is the clustering of mutants.
Our research consists of two parts in which we cluster mutants.
To cluster mutants we need a mutation testing tools.
Three mutation testing tools for Java are compared and one is selected for further use in this research.
The first part consists of a white box approach and the second part of a black box approach.
Our research uses the mutation testing tool PIT.
We identified characteristics to represent the mutants such that we can cluster them.
For our white box approach we used hierarchical clustering and for our black box approach we trained the fuzzy c-means  model.
We calculated a weighted mutation score and compared this with the mutation score of a full set of mutants executed.
Results show that with hierarchical clustering we can reduce the amount of mutants executed while maintaining the effectiveness.
The clusters generated by the machine learning model were less accurate and showed a significant decrease in effectiveness.
\end{abstract}

@@

\section{Introduction}
The high cost requirement is often a barrier for adopting mutation testing \cite{Pizzoleto2019}.
A lot of techniques and methods have been developed to improve the performance, however, most of these approaches are not as effective as mutation testing a full set of mutants\cite{Pizzoleto2019,Yao2014}. 
\newline
A technique to reduce the number of mutants or the number of test case executions is mutation clustering.
Mutation clustering aims to reduce the amount of mutants to be executed by clustering mutants\cite{Ma2016,Yu2019PossibilityScope}.
The clustering of mutants has been researched with promising results\cite{Ji2009,Wilinski2015,Ma2016}. For example Ma et al., \cite{Ma2016} clustered the mutants for expressions. 
However in their paper they indicate that the choosing expressions as the domain of their cluster is a limitation.
A recurring problem in the research of mutation clustering is to determine the means of domains as the centre of a cluster\cite{Ji2009,Wilinski2015,Ma2016,Wei2021SpectralTesting}.
\newline
To reduce the cost of mutation testing, we try to find a solution that can cluster mutants while maintaining the accuracy of the same complete set of mutants. 
Our goal is to cluster any mutant that is generated within the Java programming language.
This research aims to remove the scoping limitations present in existing research.
There should be no requirements for which mutants can be clustered.
We do this with two different techniques: qualitative and quantitative.
We devise a white and black box approach for clustering mutants.
Our white box approach contains a qualitative analysis on mutants and a methodology to cluster them.
The black box approach makes use of a machine learning model to cluster mutants.
We structure our research on the following research questions:
\newline
\textbf{Research Question 1}: What set of characteristics can we identify for clustering generated mutants to reduce the amount executed, while maintaining effectiveness?
\newline
\textbf{Research Question 1.1}: How do the existing mutation testing tools for Java compare to each other?
\newline
\textbf{Research Question 2}: How can we train a machine learning model to recognize and cluster generated mutants to reduce the amount executed, while maintaining effectiveness?

\subsection{Contributions}
Our research makes the following contributions:
\begin{enumerate}
 \item A white box methodology to cluster generated mutants based on chosen characteristics to reduce the cost of mutation testing.
  \item A black box methodology to cluster generated mutants based on chosen characteristics to reduce the cost of mutation testing.
 \item A proof of concept which implements the qualitative methodology chosen and elaborated within the thesis.
 \item A proof of concept which implements the quantitative approach to cluster mutants and is elaborated within the thesis. 
\end{enumerate}
\subsection{Outline}

@@

\section{Selecting a mutation testing tool}
We reviewed literature on three different mutation testing tools[CITE OWN PAPER HERE].
However the most recent study is from 2017.
At the time of writing this paper these studies are at least four years old. 
Some of the tools were updated or are still under active development\cite{pit-releases,Major}.
There are three candidate tools; Major, MuJava and PIT.
We extend the existing mutant operator comparison with the operators that have been added since the publishing date of the existing literature.
Additionally we add an overview of the overlapping mutant operators for the new mutant operators.

\todo{show tables here? maybe a summarized version of this?}
\todo{the results can be found in ref to paper??}

MuJava is not actively maintained and has not been updated in the last few years.
It does not support JUnit 4 and all versions of TestNG\cite{mujava}.
These are test engines and are crucial for executing tests.
MuJava also does not support source projects with java version 1.7 or higher\cite{mujava}.
Conforming to these requirements would reduce the set of projects we could use in or experiments.
\newline
While Major supports JUnit 4,
we did not succeed in generating mutants with this tool[CITE OWN PAPER HERE].
It would be too time consuming to customize all source projects to work with Major.
Furthermore there is not much documentation available for this tool.
\newline
PIT targets the industry, is actively maintained and is open source\cite{Kintis2016AnalysingStudy}.
It supports Maven, Gradle, has a command line interface and has a faster execution time than the other tools.
For example PIT provides a plugin system in which you can inject your own code in various stages of mutation testing process\cite{pit}.
PIT also generated significantly more mutants in every project we have tested[CITE OWN PAPER HERE].
Based on the information presented in this chapter we decide that PIT is the best choice for generating and executing mutants for this research.

\section{Qualitative clustering approach}
\section{Quantitative clustering approach}
\section{validation}
